---
title: "Project: Group 3"
author: "EL Ghali DEBBAGH and Victoria SEMINA Project"
format: pdf
editor: visual
execute: 
  warning: false
  error: false
  echo: false
---

```{r}
#| message: false
#| warning: false
here::i_am("project.Rproj")
library(here)
library(vroom)
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(psych)
library(lmtest)

```

## Research Question

The research question of our project: How do various development indicators, such as national income, investment, and education, interact with and influence the levels of food (in)security within a nation? It seemed to us that it would be relevant to study the relationship between food security and various indicators of the country's development, since this directly affects the overall well-being of the country and the quality of life in it. In addition, we look at both developed and developing countries, which will allow us to identify patterns of inequality. Analyzing the relationships between food security and development outcomes can also highlight inequalities in access to resources and opportunities.

It seemed to us that, in general, the study of food security and its links with various development indicators is important for the scientific community and for our own development, in order to potentially understand how to build a more sustainable, prosperous and international society.

As part of the analysis, we expect that a certain relationship will be found between national income and food security. Specifically, that high national income can be associated with increased access to quality and varied food. However, despite this, food security problems such as malnutrition due to social inequalities may also exist in developed countries.

In addition, we are going to analyze the relationship between education and food security. We expect that higher levels of education will generally be associated with higher levels of healthy eating awareness and resource management skills, which will help improve food security.

We also want to analyze developed and developing countries because we expect different results in these groups. It seems to us that in developed countries the main problems will be related to the social aspects of food security, such as inequality in access to resources. In particular, anorexia and obesity may be problems of overconsumption and pressure from social standards of beauty, while in developing countries, conversely, low levels of education and limited investment may lead to malnutrition and lack of access to key food resources. Overall, through the analysis of the relationships between food security and various development indicators, we want to identify the main dynamics that shape the level of food security in different countries.

## Data Set Description

### Food Security Indicators:

The first data source which we used is FAO (Food and Agriculture Organisation of the United Nations): https://www.fao.org/faostat/en/#data/FS with the data selected from "Suite of Food Security Indicators" (steps to get the final table are: select in countries: "select all", in elements select "values", in items select "prevalence of undernourishment (percent)", "Prevalence of severe food insecurity in the total population (percent)", "Percentage of children under 5 years affected by wasting (percent)", "Percentage of children under 5 years of age who are stunted (modelled estimates) (percent)", "Percentage of children under 5 years of age who are overweight (modelled estimates) (percent)", "Prevalence of obesity in the adult population (18 years and older)", "Prevalence of anemia among women of reproductive age (15-49 years)", "Prevalence of low birthweight (percent)", "Prevalence of exclusive breastfeeding among infants 0-5 months of age", for the selected years , we choose to work from 2013 to 2022.

We decided to choose these indicators because, apart from basic nutrition, food security is linked to economic stability, long-term health, women's empowerment and the environment. Research reveals that food insecurity can result in ongoing short and long-term health problems. The first two years of a child's life are of critical importance for their healthy growth and development. For children, severe food insecurity has been linked to chronic health conditions like stunting, wasting and anemia. A diet lacking enough calories, protein, vitamins and minerals will impede a child's growth and development from before they're born up until adolescence. A pregnant mother's malnutration can lead to low birthweight, infant mortality, preterm delivery and slow cognitive development for her baby.

All in all, this suite of food security indicators is composed of a revised selection of indicators based on the SDGs monitoring requirement and data availability. For example, some of the utilization indicators are replaced with official SDGs indicators. Data in this domain are featured in the State of Food Security and Nutrition in the World 2022 report.

```{r}
#| message: false

FAO <- vroom("FAO.csv")
```

```{r}
#First of all, we had a problem with the FAO database because some years in column 'Year' were written as an interval, e.g. 2014 was written as 2013-2015. This problem actually comes from the original FAO website's database description, where years are written as following 2022/2021-2023. So first of all we figured out this issue.


FAO$Year <- ifelse(grepl("\\d{4}-(\\d{4})", FAO$Year), sub("(\\d{4})-(\\d{4})", "\\1", FAO$Year) %>% as.numeric() + 1, FAO$Year)

```

```{r}
# Then we had a need to delete the 'Year code column' since we already have a column year
FAO_test<-
FAO |> select(-`Year Code`)


```

```{r}
#After that, we use 'pivot wider' to put our indicators in columns to get our cleaned first table
FAO_widetest<- FAO_test|> pivot_wider(names_from = Item, values_from = Value, id_cols =c('Year', 'Area'))
```

### World Development Indicators

The second data source is The World Bank where we got the world development indicators (https://databank.worldbank.org/source/world-development-indicators#). To get our final table, we select in database "World Development Indicators", for Countries, we selected all countries from Afghanistan to Zimbabwe, in series we select "GDP growth (annual %)", "GDP per capita growth (annual %)", "Agricultural land (% of land area", "Children in employment, female (% of female children ages 7-14)", "Children in employment, female (% of female children ages 7-14)", "Children in employment, male (% of male children ages 7-14)", "Children out of school (% of primary school age)", "Children out of school, female (% of female primary school age)", "Children out of school, male (% of male primary school age)", "Death rate, crude (per 1,000 people)", "Current education expenditure, total (% of total expenditure in public institutions)", "Current health expenditure (% of GDP)", "Food production index (2014-2016 = 100)","Gini index" ,"Labor force participation rate, female (% of female population ages 15+ (modeled ILO estimate)", "Labor force, female (% of total labor force)", "Life expectancy at birth, total (years)", "Lifetime risk of maternal death (%)", "Number of infant deaths".

World Development Indicators (WDI) is the primary World Bank collection of development indicators, compiled from officially recognized international sources. It presents the most current and accurate global development data available, and includes national, regional and global estimates. Even though Global Development Finance (GDF) is no longer listed in the WDI database name, all external debt and financial flows data continue to be included in WDI. The GDF publication has been renamed International Debt Statistics (IDS), and has its own separate database, as well. The World Bank is known for maintaining high standards in databases. There is data reliability, World Bank data is generally subject to rigorous standards and quality controls. As part of our project, the database from World Bank enable to make international comparisons by collecting data from many countries.

The chosen indicators seem quite obvious; those in the first table mainly concern the percentages of people suffering from diseases, such as anemia, obesity, low birth weight, for the second table the indicators are as common as the first; these are the basic development indicators (GDP and GDP/inhabitant, etc.) as well as indicators on education, in particular the education of women, then indicators on resources and agriculture going to indicators of inequalities like the Gini index.

```{r}
#| message: false

WB<- vroom("WB.csv")
```

```{r}
# Now we are looking at the table from the World Bank. The initial problem there was the fact that years were written in columns instead of lines. That's why we decided to use 'pivot longer' to put years in lines (to combine after with the first table)
WB_longtest<-WB|>pivot_longer(!(`Country Name` | `Country Code` |`Series Name` |`Series Code`))
```

```{r}
#Here we need to name the new column with years as 'Year'
WB_longtest <- WB_longtest |> rename("Year"="name")
```

```{r}
# In the World Bank database some missing values were automatically replaced by ".." In order not to get any problems with working with tables we replace such missed values with NA
WB_longtest[WB_longtest==".."]<- NA
```

```{r}
# After cleaning this, we faced with the another problem with  year columns where years were written as 2014[YR2014]. However, we need to have only the year insteasm so we deleted these brackets for every year 
WB_longtest$Year <- gsub("\\[.*\\]", "", WB_longtest$Year)
```

```{r}
#After that, we use 'pivot wider' to put our indicators in columns to get our cleaned second table
WB_wide <- WB_longtest |>
  pivot_wider(
    names_from = `Series Name`,
    values_from = value,
    id_cols = c('Year', 'Country Name'),
    values_fn = list
  )

```

```{r}
#| message: false

#Summary of the first table
num_rows <- nrow(FAO_widetest)
num_cols <- ncol(FAO_widetest)

missing_values <- colSums(is.na(FAO_widetest))

summary_table <- data.frame(
  Number_of_rows = num_rows,
  Number_of_columns = num_cols,
  Number_of_missing_values = missing_values
)


knitr::kable(summary_table)
```

```{r}
#Summary of the second table

num_rows_wide <- nrow(WB_wide)
num_cols_wide <- ncol(WB_wide)


missing_values_wide <- colSums(is.na(WB_wide))


summary_table_wide <- data.frame(
  Number_of_rows = num_rows_wide,
 Number_of_columns = num_cols_wide,
  Number_of_missing_values = missing_values_wide
)


knitr::kable(summary_table_wide)

```

```{r}
WB_wide <- WB_wide |> 
  rename("Area" = `Country Name` )
```

```{r}
WB_wide <- slice(WB_wide, 1:2170)

```

```{r}
data <- merge(WB_wide, FAO_widetest, by = c("Year", "Area"))
```

```{r}
WB_wide$Year <- trimws(WB_wide$Year)
WB_wide$Area <- trimws(WB_wide$Area)

FAO_widetest$Year <- trimws(FAO_widetest$Year)
FAO_widetest$Area <- trimws(FAO_widetest$Area)

```

```{r}
data <- merge(WB_wide, FAO_widetest, by = c("Year", "Area"), all.x = TRUE, all.y = TRUE)

```

```{r}
#| warning: false


area_column <- data$Area

non_numeric_columns <- sapply(data, function(x) !is.numeric(x))


data[, non_numeric_columns] <- lapply(data[, non_numeric_columns], function(x) as.numeric(as.character(x)))

data <- round(data, 2)

data$Area <- area_column
```

## Data Analysis

### First part

First of all, let's compare GDP and Prevalence of low birthweight:

```{r}
data_plot <- data

year_least_missing <- which.min(apply(data_plot[, c("GDP growth (annual %)", "Prevalence of low birthweight (percent)")], 2, function(x) sum(is.na(x))))
year_least_missing <- unique(data_plot$Year[year_least_missing])

data_plot <- na.omit(data_plot[data_plot$Year == year_least_missing, c("Year", "Area", "GDP growth (annual %)", "Prevalence of low birthweight (percent)")])

data_plot$Prevalence_Category <- cut(data_plot$`Prevalence of low birthweight (percent)`,
                                      breaks = quantile(data_plot$`Prevalence of low birthweight (percent)`, probs = seq(0, 1, 0.2)),
                                      labels = c("Très faible", "Faible", "Moyenne", "Élevée", "Très élevée"),
                                      include.lowest = TRUE)

ggplot(data = data_plot, aes(x = `GDP growth (annual %)`, y = `Prevalence of low birthweight (percent)`, color = Prevalence_Category)) +
  geom_point(size = 3) +
  labs(x = "GDP growth (annual %)", y = "Prevalence of low birthweight (percent)",
       title = paste("Relation entre GDP growth et Prevalence of low birthweight pour l'année", year_least_missing),
       color = "Catégorie de Prévalence") +
  theme_minimal() +
  theme(legend.position = "bottom")  # Positionner la légende en ba
```

```{r}


data_plot <- data

year_least_missing <- which.min(apply(data_plot[, c("GDP growth (annual %)", "Prevalence of low birthweight (percent)")], 2, function(x) sum(is.na(x))))
year_least_missing <- unique(data_plot$Year[year_least_missing])

data_plot <- na.omit(data_plot[data_plot$Year == year_least_missing, c("Year", "Area", "GDP growth (annual %)", "Prevalence of low birthweight (percent)")])

correlation_matrix <- cor(data_plot[, c("GDP growth (annual %)", "Prevalence of low birthweight (percent)")])

summary(data_plot[, c("GDP growth (annual %)", "Prevalence of low birthweight (percent)")])

print(correlation_matrix)
```

The values we got in the correlation matrix seem to indicate a very weak correlation (close to zero) between "GDP growth (annual %)" and "Prevalence of low birthweight (percent)".

The correlation of 1 on the diagonal represents the perfect correlation between each variable with itself, which is expected.

However, the non-diagonal values (0.09525624) suggest a very weak correlation between the two variables. This low correlation may explain why the scatter plot does not show a clear trend.

```{r}
data_plot_2020 <- data

data_plot_2020 <- na.omit(data_plot_2020[data_plot_2020$Year == 2020, c("Year", "Area", "GDP growth (annual %)", "Prevalence of low birthweight (percent)")])

data_plot_2020$Prevalence_Category <- cut(data_plot_2020$`Prevalence of low birthweight (percent)`,
                                          breaks = quantile(data_plot_2020$`Prevalence of low birthweight (percent)`, probs = seq(0, 1, 0.2)),
                                          labels = c("Très faible", "Faible", "Moyenne", "Élevée", "Très élevée"),
                                          include.lowest = TRUE)

correlation_matrix_2020 <- cor(data_plot_2020[, c("GDP growth (annual %)", "Prevalence of low birthweight (percent)")])

summary(data_plot_2020[, c("GDP growth (annual %)", "Prevalence of low birthweight (percent)")])

print(correlation_matrix_2020)

ggplot(data = data_plot_2020, aes(x = `GDP growth (annual %)`, y = `Prevalence of low birthweight (percent)`, color = Prevalence_Category)) +
  geom_point(size = 3) +
  labs(x = "GDP growth (annual %)", y = "Prevalence of low birthweight (percent)",
       title = "Relation entre GDP growth et Prevalence of low birthweight pour l'année 2020",
       color = "Catégorie de Prévalence") +
  theme_minimal() +
  theme(legend.position = "bottom")  # Positionner la légende en bas


```

The correlation coefficients between "GDP growth (annual %)" and "Prevalence of low birthweight (percent)" for the year 2020 are 0.07955535. This indicates a very weak correlation between these two variables for this specific year.

The close to zero correlation suggests that there is no strong linear relationship between "GDP growth (annual %)" and "Prevalence of low birthweight (percent)" in 2020.

In the analysis of the dataset, two key years were examined to explore the relationship between GDP growth and the prevalence of low birthweight -- 2013 and 2020. The choice of these years was guided by the availability of data, aiming to minimize missing values.

In 2013, the correlation analysis between GDP growth and the prevalence of low birthweight did not reveal a clear trend. The correlation coefficient was close to zero, indicating a weak linear relationship between these two indicators for this year.

Moving to the year 2020, similar findings were observed. The correlation coefficient remained low, suggesting a limited linear association between GDP growth and the prevalence of low birthweight.

These results imply that, at least within the confines of linear correlation, there is no strong apparent connection between GDP growth and the prevalence of low birthweight for the selected years. Further exploration and consideration of additional factors or non-linear relationships may provide a more comprehensive understanding of the dynamics between economic development and health indicators.

Moving to another indicator: "Prevalence of severe food insecurity in the total population (percent) (3-year average)"

```{r}
#| warning: false

selected_cols1 <- c("Year", "Area", "GDP growth (annual %)", "Prevalence of severe food insecurity in the total population (percent) (3-year average)")

selected_data1 <- data[, selected_cols1]

filtered_data1 <- selected_data1[complete.cases(selected_data1[, "Prevalence of severe food insecurity in the total population (percent) (3-year average)"]), ]

year_with_least_na <- which.min(table(filtered_data1$Year))

data_for_plot1 <- filtered_data1[filtered_data1$Year == unique(filtered_data1$Year)[year_with_least_na], ]

ggplot(data = data_for_plot1, aes(x = `GDP growth (annual %)`, y = `Prevalence of severe food insecurity in the total population (percent) (3-year average)`)) +
  geom_point(size = 3, color = "blue") +
  labs(x = "GDP growth (annual %)", y = "Prevalence of severe food insecurity (percent) (3-year average)",
       title = paste("Relation entre GDP growth et Prevalence of severe food insecurity (", unique(data_for_plot1$Year), ")")) +
  theme_minimal()


```

```{r}
selected_cols2 <- c("GDP growth (annual %)", "Prevalence of severe food insecurity in the total population (percent) (3-year average)")

selected_data2 <- data[, selected_cols2]

selected_data2 <- na.omit(selected_data2)

correlation_table2 <- cor(selected_data2)

print(correlation_table2)

```

The results of the correlation between "GDP growth (annual %)" and "Prevalence of severe food insecurity in the total population (percent) (3-year average)" for the available data are close to zero, indicating a very weak correlation. The correlation is practically zero (-0.0016), suggesting that there is no obvious linear relationship between these two variables.

This means that, at least based on the linear correlation, there is no clear trend indicating that the GDP growth rate is associated with the prevalence of severe food insecurity in the population.

```{r}
filtered_data_non_na <- filtered_data1[complete.cases(filtered_data1[, c("GDP growth (annual %)", "Prevalence of severe food insecurity in the total population (percent) (3-year average)")]), ]

second_year_with_least_na <- which.min(table(filtered_data_non_na$Year)[-year_with_least_na])

second_data_for_plot <- filtered_data_non_na[filtered_data_non_na$Year == unique(filtered_data_non_na$Year)[second_year_with_least_na], ]

correlation_second_year <- cor(second_data_for_plot[, c("GDP growth (annual %)", "Prevalence of severe food insecurity in the total population (percent) (3-year average)")])

print(correlation_second_year)


```

This correlation is weak and negative, indicating a weak linear relationship between these two variables for this specific year. This means that there is little evidence of a linear relationship between the GDP growth rate and the prevalence of severe food insecurity in the population for that particular year.

As always, it is important to note that correlation only captures linear relationships and does not guarantee causation.

```{r}
ggplot(data = second_data_for_plot, aes(x = `GDP growth (annual %)`, y = `Prevalence of severe food insecurity in the total population (percent) (3-year average)`)) +
  geom_point(size = 3, color = "blue") +
  labs(x = "GDP growth (annual %)", y = "Prevalence of severe food insecurity (percent) (3-year average)",
       title = paste("Relation entre GDP growth et Prevalence of severe food insecurity (", unique(second_data_for_plot$Year), ")")) +
  theme_minimal()

```

### Conclusion of the first part:

In our analysis, we examined the correlation between GDP growth and two key indicators of food insecurity -- the prevalence of severe food insecurity and the prevalence of low birthweight. The results consistently showed weak correlations between GDP growth and both indicators for the selected years, suggesting a limited linear relationship between economic development and these dimensions of food insecurity.

The correlation coefficients were close to zero, indicating that changes in GDP growth are not strongly associated with variations in the prevalence of severe food insecurity or low birthweight. This observation challenges the assumption that economic growth alone is sufficient to alleviate these aspects of food insecurity.

These findings prompt further investigation into alternative development indicators that may offer a more comprehensive understanding of the relationship between food insecurity and a nation's overall development. It is plausible that other factors, such as educational attainment, healthcare infrastructure, or social policies, could play crucial roles.

As we move forward, considering additional indicators and employing a more nuanced approach may help unveil hidden patterns and provide a clearer picture of the intricate interplay between development and food security. Moreover, assessing the disparities among countries, particularly focusing on the socio-economic conditions of the least developed nations, could offer valuable insights into addressing food insecurity on a global scale.

This ongoing exploration is vital not only for understanding the multifaceted nature of the issue but also for formulating effective policies and interventions that address the root causes of food insecurity, contributing to a more equitable and sustainable world.

### Second part

We further decided to conduct the analysis using the variables Food Production Index and Prevalence of Malnutrition because we believe that these are important for assessing the efficiency of the agricultural system and its impact on the food security of the society.

We believe that analysis of the Food Production Index and the Prevalence of Malnutrition is important for an overall assessment of the sustainability of a country's food system. If the index shows a high level of food production, but the prevalence of undernourishment is still high, this may indicate problems in food distribution, access to food, or other factors preventing the population from effectively feeding.

Globally, for researchers, comparing food production data with malnutrition rates can reveal problems in the supply chain, such as product wastage, poor storage and transportation efficiency, and problems in trade and distribution. Among other things, for international organizations this analysis may be especially important, as we think, since it helps to identify those areas of activity that require improvement. For example, if agricultural output is high but the prevalence of malnutrition is also high, measures may be needed to improve population access to food, nutrition education, and nutrition.

In terms of our project Food Production Index and Prevalence of Malnutrition helps understand the relationship between food production and food availability.

```{r}

selected_year <- 2013
selected_vars <- c("Area", "Food production index (2014-2016 = 100)", "Prevalence of undernourishment (percent) (3-year average)")
scatter_data <- data[data$Year == selected_year, selected_vars]

scatter_data <- na.omit(scatter_data)



ggplot(scatter_data, aes(x = `Food production index (2014-2016 = 100)`, y = `Prevalence of undernourishment (percent) (3-year average)`)) +
  geom_point(color = "orange") +
  labs(title = paste("Scatter Plot of Food Production Index vs. Prevalence of Undernourishment (2013)"),
       x = "Food Production Index (2014-2016 = 100)",
       y = "Prevalence of Undernourishment (percent)") +
  theme_minimal()



```

```{r}

selected_year <- 2013
selected_vars3 <- c("Food production index (2014-2016 = 100)", "Prevalence of undernourishment (percent) (3-year average)")
correlation_data3 <- data[data$Year == selected_year, selected_vars3]

correlation_data3 <- na.omit(correlation_data3)

correlation_coefficient3 <- cor(correlation_data3$`Food production index (2014-2016 = 100)`, correlation_data3$`Prevalence of undernourishment (percent) (3-year average)`)

cat("Correlation Coefficient:", correlation_coefficient3, "\n")

```

A correlation coefficient of 0.237 indicates a positive correlation between "Food production index" and "Prevalence of undernourishment" for the year 2013. While the correlation is positive, it's not very strong.

Now let's have a look at the same variables but in 2018.

```{r}

year_2018 <- 2018
vars_2018 <- c("Food production index (2014-2016 = 100)", "Prevalence of undernourishment (percent) (3-year average)")
correlation_data_2018 <- data[data$Year == year_2018, vars_2018]

correlation_data_2018 <- na.omit(correlation_data_2018)

correlation_coefficient_2018 <- cor(correlation_data_2018$`Food production index (2014-2016 = 100)`, correlation_data_2018$`Prevalence of undernourishment (percent) (3-year average)`)

cat("Correlation Coefficient for 2018:", correlation_coefficient_2018, "\n")

```

The examination of the correlation between Food Production Index and the Prevalence of Undernourishment for these years reveals a notable pattern --- the correlation coefficients for both 2013 (0.237) and 2018 (-0.155) are close to zero. This near-zero correlation suggests a weak and possibly non-existent linear relationship between these two indicators.

The proximity to zero in correlation coefficients implies that variations in food production do not consistently or strongly predict changes in the prevalence of undernourishment across countries.

So, we can say that the weak correlation underscores the complexity of factors influencing food security, indicating that food production alone may not be a decisive factor in determining the prevalence of undernourishment.

### Conclusion of the second part:

In general, the fact that the connection turned out to be insignificant is quite intuitive.

Firstly, in the era of globalization, the production of products in one country does not mean at all that they will be used for consumption in that same country.

Poor countries often produce goods for export and starve themselves, while northern countries rich in natural resources and human capital do not have a climate suitable for agriculture, but can nevertheless easily buy these products for consumption by their citizens.

In addition, the Food Production Index is a fairly complex variable and often includes specific food products, but does not, for example, include the production of fresh water or other essential nutrients.

## Pushing the analysis

Previous analysis didn't show us any clear tendency on how the economic level influences food insecurity.

So, before analyzing socio-economic indicators, such as education and woman role in the society, we decided to try to push a little more our analysis.

First of all, we provided a table with 5 richest and 5 poorest countries in year 2013 by PIB level (source FMI) and focus on specific indicator which is "Prevalence of anemia among women of reproductive age (15-49 years)" to try to figure out if any influence of the GDP level (not the growth, because, as know, GDP grows faster in developping countries rather than developped) on this indicator.

The five richest countries according to PIB level in 2013 are: USA, China, Japan, Germany, France. The five poorest countries in 2013 according to same source and on the same criteria are Syria, Tuvalu, Nauru, Kiribati and Marshall islands.

```{r}
selected_countries <- c("United States", "China", "Japan", "Germany", "France", "Syrian Arab Republic", "Tuvalu", "Nauru", "Kiribati", "Marshall Islands")

data_countries_PIB <- data[data$Area %in% selected_countries, ]



```

```{r}
#| warning: false



graphPIB <- ggplot(data_countries_PIB, aes(x = Year, y = `Prevalence of anemia among women of reproductive age (15-49 years)`, group = Area, color = Area)) +
  geom_line() +
  labs(title = "Evolution of Anemia Prevalence by Country (2013-2019)",
       x = "Year",
       y = "Prevalence of Anemia (%)",
       color = "Country") +
  scale_x_continuous(breaks = seq(2013, 2019, 1)) +  # Sélectionner les années de 2013 à 2019
  theme_minimal()

print(graphPIB)


```

### Difference between poorest and richest countries

We decided to demonstrate the prevalence of anemia on a graph, since this disease is one of the key diseases in the world, including having a special impact during pregnancy and directly affecting the development of the unborn child.

At the top of the graph we can see the poor countries with the highest rates of anemia, compared to the richest countries further down the graph.

Of particular note is the fact that poor countries have significant data problems, often complete lack of data, as can be seen in our data_countries_PIB table, in which several indicators of food insecurity are almost empty for the poorest countries.

## Regression analysis

Of course, indicators demonstrating the level of food are important, but as our previous analysis showed, we still have a need to add other socio-economic indicators.

Therefore, we added to the regression "Female labor force participation rate (% of female population aged 15 years and over) (modeled ILO estimate)", "Current education expenditure, total (% of total expenditure in public institutions"), and See also "Children out of school (% of primary school age)".

We took the following indicators because we believe that they can directly influence the standard of living, and therefore the nutrition of the population.

In particular, if children do not attend school from an early age, this indicates that they begin to work from an early age. As a rule, this is low-skilled and low-paid work without further possibility of any career growth and retraining in other specialties.

```{r}
complete_data <- data[complete.cases(data$`Prevalence of undernourishment (percent) (3-year average)`, data$`Labor force participation rate, female (% of female population ages 15+) (modeled ILO estimate)`), ]

cor(complete_data$`Prevalence of undernourishment (percent) (3-year average)`, complete_data$`Labor force participation rate, female (% of female population ages 15+) (modeled ILO estimate)`)

```

```{r}

formula <- formula(`Prevalence of undernourishment (percent) (3-year average)` ~ 
                    `Labor force participation rate, female (% of female population ages 15+) (modeled ILO estimate)` + 
                    `Current education expenditure, total (% of total expenditure in public institutions)` + 
                    `Children out of school (% of primary school age)`)
                    
model <- lm(formula, data = data)

summary(model)


```

So, we made regression analysis and we have tre following results now.

The intercept is 14.32. This is the estimated value of the prevalence of undernourishment when all explanatory variables are equal to zero.

Labor force participation rate, female (%) (modeled ILO estimate): The coefficient is 0.04301 with a p-value of 0.4009. This indicates that for a one unit increase in the female labor force participation rate, the prevalence of undernourishment increases by 0.04301, but this increase is not statistically significant.

Current education expenditure, total (% of total expenditure in public institutions): The coefficient is -0.09698 with a p-value of 0.1546. This suggests that a one-unit increase in current education spending, as a percentage of total spending in public institutions, is associated with a 0.09698 decrease in the prevalence of undernourishment, but this decrease is not not statistically significant.

Children out of school (% of primary school age): The coefficient is 0.07915 with a p-value of 0.1093. This means that a one unit increase in the percentage of children out of school among those of school age is associated with an increase of 0.07915 in the prevalence of undernourishment, but this increase is not statistically significant.

R-squared: The coefficient of determination (R²) is 0.03776, which suggests that the model explains only about 3.78% of the variance in the prevalence of undernourishment.

We can say again that the results do not show significant relationships between the explanatory variables and the prevalence of undernourishment. High p-values indicate that the coefficients are not statistically different from zero.

So, it is possible that other variables not included in the model could have a more significant influence on the prevalence of undernutrition, that's why we decided to create a formula with all the World Bank indicators.

```{r}
selected_indicators <- c("GDP growth (annual %)", "Agricultural land (% of land area)", "Current education expenditure, total (% of total expenditure in public institutions)", "Life expectancy at birth, total (years)")

selected_indicators <- paste("`", selected_indicators, "`", sep = "")

formula_selected <- as.formula(paste("`Prevalence of undernourishment (percent) (3-year average)` ~", paste(selected_indicators, collapse = " + ")))

model_selected <- lm(formula_selected, data = data)

summary(model_selected)

```
